Evolutionary rescue and reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

### 2023-05-17 doc/manuscript/translocation.qmd
Minor edits in response to TS & AB review

<!--# Intended outlet is PNAS. Preferred length of research reports is 6 pages, but up to 12 pages is allowed. A standard 6-page article is approximately 4,000 words, 50 references, and 4 medium-size graphical elements (i.e., figures and tables) -->

@@

<!--# Title should be less than 135 characters -->

@@

Roland A. Knapp^1,2,^\*, Mark Q. Wilber^3,a^, Allison Q. Byrne^4,5,a^, Maxwell B. Joseph^6,7^, Thomas C. Smith^1,2^, Andrew P. Rothstein^4,5^, Rob L. Grasso^8^, Erica Bree Rosenblum^4,5^

@@

^1^Sierra Nevada Aquatic Research Laboratory, University of California, Mammoth Lakes, CA, 93546

^2^Earth Research Institute, University of California, Santa Barbara, CA, 93106-3060

^3^School of Natural Resources, University of Tennessee Institute of Agriculture, Knoxville, TN, 37996

^4^Department of Environmental Science, Policy, and Management, University of California - Berkeley, Berkeley, CA, 94720-3114

^5^Museum of Vertebrate Zoology, University of California - Berkeley, Berkeley, CA, 94720-3160

^6^Earth Lab, University of Colorado, Boulder, CO, 80303

^7^Planet, San Francisco, CA, 94107

^8^Resources Management and Science, Yosemite National Park, El Portal, CA, 95318

@@

^a^These authors contributed equally to this work.

**Preprint Server:** bioRxiv

@@

@@

Understanding how species persist despite accelerating global change is critical for the conservation of biodiversity.
One possibility is rapid evolution that restores positive growth to declining populations ("evolutionary rescue").
Individuals from these rescued populations could also be used to reestablish exinct populations.
We evaluated both possibilities in a once-common frog species driven to the brink of extinction by a disease that has devastated global amphibian biodiversity and for which few options exist to reverse disease impacts.
We found that frogs have evolved increased resistance to infection, facilitating population recovery.
In addition, introduction of frogs from these populations allowed reestablishment despite the ongoing presence of disease.
Such reintroductions could play an important role in biodiversity conservation in our rapidly changing world.

@@

Human activities are intensifying threats to global biodiversity [@ceballos2015], with important implications for ecosystem resilience and human welfare [@naeem2009].
To persist, affected taxa will often need to adapt to their altered environment or face decline and extinction.
"Evolutionary rescue" describes rapid evolutionary change that increases the frequency of adaptive alleles and thereby restores positive population growth [@carlson2014].
In addition to preventing extinction of declining populations, this process could also result in the seeding of nearby empty habitats and allow population reestablishment and recovery.
Although this overall scenario is potentially of critical importance for the conservation of an ever-growing number of threatened taxa, descriptions of its operation in the wild are lacking.

One consequence of human alteration of the biosphere is an increase in emerging infectious diseases [@jones2008; @fisher2012].
Such diseases pose a severe threat to wildlife populations [@daszak2000], and have caused dramatic declines and extinctions in a wide range of taxa [@hewson2014; @samuel2015; @scheele2019; @cunningham2021].
Amid this global loss of biodiverity, some populations decimated by disease have subsequently recovered to varying extents [@newell2013; @voyles2018; @knapp2016].
Although we know relatively little about the mechanisms underlying recovery [@brannelly2021; @russell2020], host evolution is one possibility [@carlson2014; @searle2020].
This process is known from a diversity of wildlife populations [@savage2016; @epstein2016; @gignoux-wolfsohn2021; @holland2022], and often results in increased host resistance and/or tolerance (see [@medzhitov2012] for definitions).
However, linking these evolutionary responses to population recoveries is difficult and few clear examples exist.
In addition to promoting recovery of disease-impacted populations, the resistant/tolerant individuals in such populations could potentially be used to reestablish populations that were extirpated following disease emergence [@mendelson2019; @brannelly2021].
To date, the use of resistant/tolerant individuals for species recovery is rare [but see @joseph2018], and studies that evaluate its potential utility are critically needed.

@@

In the current study, we combine results from genomic analyses, a 15-year landscape-scale frog reintroduction effort, and a detailed population model to provide an unprecedented example of how evolution in response to disease allows the successful reestablishment of extirpated MYL frog populations.
Specifically, we used exome capture methods to show that MYL frogs have undergone substantial evolutionary change associated with Bd exposure and that might provide a genetic basis for the increased resistance/tolerance to Bd infection observed in naturally recovering populations [@knapp2016].
We then use the reintroduction of resistant/tolerant frogs to 24 sites and results from long-term capture-mark-recapture (CMR) surveys to describe the dynamics and drivers of population establishment.
These data are also used to parameterize a population model that extends our inferences of population establishment well beyond the temporal extent of the CMR data.
Collectively, these results provide one of the first examples of how evolutionary rescue and the reintroduction of resistant/tolerant animals can allow the large-scale recovery of disease-impacted species.

@@

To determine whether MYL frog populations show an evolutionary response to Bd, we compared frog genomes between populations with contrasting histories of Bd exposure.
Specifically, we compared frog genomes sampled in four populations that have not yet experienced a Bd-caused epizootic ("naive") [@zhou2015] versus in five populations that experienced a Bd epizootic during the past several decades and have since recovered to varying degrees ("recovering"; @fig-allelefrequencies) [@knapp2016; @vredenburg2010].
These recovering populations exist in an enzootic state [@briggs2010], characterized by high Bd prevalence and, in adults, moderate Bd infection intensities ("load") that are typically well below the level expected to cause mortality [@vredenburg2010].
(See SI for additional details regarding the study design.)

@@

Results from the individual variant and splined window analyses show that recovering populations have signatures of strong selection on multiple regions of the genome.
The analysis of individual variants identified 11 "outlier" SNPs (i.e., showing significantly different allele frequencies between naive versus recovering populations) from 7 distinct genes across 4 contigs (@fig-selectionresults B, C).
One of the 7 identified genes did not have an associated annotation (LOC108802036<!--# Allie - check for accuracy -->).
For the outlier SNPs, frequency differences between the naive and recovering populations ranged from 41% to 86%.
Most of these SNPs showed frequency differences in only a subset of the sampled populations (@fig-allelefrequencies A, B), but the SNP in the RIN3 gene showed consistent differences in frequencies across all populations (@fig-allelefrequencies C).
This is suggestive of strong - and parallel - selection at this locus associated with exposure of frogs to Bd across multiple populations.
(See SI for additional description of these results.)

The splined window analysis identified 33 outlier regions for $\pi_{diff}$ and 58 outlier regions for *F~ST~* (@fig-spline-manhattan A, B).
Of these, 9 regions were outliers for both metrics ("shared regions") and 2 of these shared regions also contained one or more outlier SNPs (described above).
A total of 33 annotated genes were found in the 9 shared regions (SI).
The largest $\pi_{diff}$, indicative of directional selection, occurred in a 163kb region on Contig19, 12.9Mb upstream of the RIN3 outlier SNP (@fig-spline-manhattan C).
This region contains approximately 500 SNPs and one annotated gene called "interferon-induced very large GTPase 1-like" (GVINP1).
Additionally, a shared outlier region on Contig1 contained two complement factor genes (C6 and C7).
Interestingly, this region had a large negative $\pi_{diff}$ value, consistent with balancing selection.
Finally, one shared outlier region on Contig8 contained one outlier SNP (TCF19) and was within 360kb of another outlier SNP (VARS) (@fig-spline-manhattan D, @fig-synteny-plot).
This region (854kb from the beginning of the outlier window to the VARS SNP) contained a total of 8 annotated genes.
In *Xenopus*, five of these genes occur in the extended major histocompatibility complex (MHC) Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the extended MHC Class III region (HSP70, LSM2, VARS) [@ohta2006].
Therefore, this region under selection is part of the extended MHC Class I and III complex and shows synteny with other amphibian genomes.

@@

Estimates of frog survival during the 1 year following translocation indicate that survival was highly variable between recipient sites, but relatively constant within recipient sites (for the subset of sites that received multiple translocations; @fig-translocation-survival).
As such, the average survival of frogs in the first translocation to a site was an important predictor of survival of frogs in subsequent translocations to that site (@fig-transsurvival-postdens).
This suggests an important effect of site characteristics and/or (site \* frog) interactions on frog survival.
In addition, 1-year survival was higher for frogs translocated later in the study period than earlier (@tbl-survival-earlylate), which we suggest resulted primarily from our improved ability over time to choose recipient sites with higher habitat quality for *R. sierrae*.
Finally, Bd loads were fairly consistent before versus after translocation and loads were nearly always well below the level indicative of severe chytridiomycosis and associated frog mortality (@fig-bdload-beforeafter, see SI text for additional details) [@joseph2018; @vredenburg2010].

@@

Disease-induced population declines are decimating global biodiversity [@daszak2000], but strategies to recover affected species are generally lacking.
Here, we tested the possibility that populations of resistant/tolerant individuals could be used to reestablish extirpated populations.
Although previously proposed [@mendelson2019; @scheele2021], this strategy has rarely been successfully implemented (see [@joseph2018] for a small-scale example).
In particular, although numerous studies have documented signatures of selection in hosts following disease-induced declines [e.g., @epstein2016; @gignoux-wolfsohn2021], we are unaware of any that used individuals bearing the resulting adaptive alleles to reestablish extirpated populations.
The results of our study provide evidence that (1) MYL frogs have experienced strong selection following exposure to a recently-emerged lethal fungal disease and that some of the genomic regions under selection are associated with genes related to disease resistance/tolerance, and (2) reintroduction of individuals from these resistant/tolerant populations allows the successful reestablishment of viable frog populations despite ongoing disease.

@@

Results from our genomic analyses strongly suggest that natural selection for adaptive alleles and resulting increased resistance/tolerance is at least partially responsible for the reduced susceptibility of frogs in recovering populations.
We identified specific alleles and genomic regions showing signatures of selection between adjacent Bd-naive and recovering MYL frog populations, consistent with strong selection following Bd exposure.
Importantly, some regions under selection are associated with cellular and immunological mechanisms known to contribute to disease resistance, including in amphibians [@zamudio2020].
For example, the MHC plays an important role in immunity.
In our study, we identified a region that shows evidence of selection in recovering populations and contains eight genes associated with either the MHC Class I or Class III regions.
These results corroborate numerous previous studies linking MHC genes to amphibian resistance/tolerance against Bd [e.g., @savage2011; @bataille2015].
Similarly, the region with the strongest indication of directional selection (as measured by $\pi_{diff}$) contains the gene GVINP1.
Several previous studies have found this gene to be differentially expressed during Bd infection [e.g., @grogan2018; @ellison2014] and in populations differing in Bd susceptibility [@grogan2018].
This gene is also strongly linked to disease in salmon, explaining a notable 20% of the resistance phenotype [@robledo2020; @robledo2018].
We also identified a region, characterized by high *F~ST~* and low $\pi_{diff}$, that contained the complement genes C6 and C7.
This could indicate that balancing selection is acting in this region of the genome to favor a diverse set of alleles, as is known for C6 in humans [@soejima2005].
Finally, the RIN3 gene showed a consistent pattern of allele frequency differences across all nine of the frog populations sampled in this study, indicating consistent selection in populations distributed across a wide geographic area.
This gene is associated with inflammation response and in *Xenopus* is highly expressed in wound stumps following amputation of tadpole tails [@fukazawa2009].

@@

To visualize the genomic relationship of our populations we conducted a PCA using the glPCA function in the adegenet R package [@jombart2008].
To detect regions of the genome under selection between naive and recovering populations, we used two approaches: (1) a multivariate linear mixed model to evaluate individual variants (SNPs and INDELs), and (2) a splined window analysis to evaluate larger genomic regions.
For the variant analysis, we first used a stringent data filter to include only variants with \< 5% missing data (missing for no more than 2 individuals), and then calculated the likelihood ratio statistic for the resulting set of 148,307 high quality variants across 127 contigs using GEMMA [@zhou2014].
GEMMA calculates and incorporates a relatedness matrix for input samples, allowing us to account for relatedness and population structure when calculating likelihood ratio statistics.
We identified variants showing different allele frequencies between naive versus recovering populations ("outliers") using a Bonferroni- corrected significance level of 0.01.
We visualized the results using a Manhattan plot and qqplot.
We developed a more liberal set of outlier variants using a Bonferroni- corrected significance level of 0.05 and used this set solely for the gene ontology (GO) analysis (see below and SI: dataset S1, S2).

@@

In the splined window analysis, we identified outlier regions using *F~ST~* and differences in nucleotide diversity ($\pi_{diff}$) between naive and recovering populations.
First, we calculated per-site *F~ST~* between the naive and recovering individuals for all bi-allelic SNPs in the 30 largest contigs (98% of all SNPs) using VCFtools [@danecek2011].
Next, we calculated per-site nucleotide diversity $\pi$ separately for individuals from the naive and recovering populations using VCFtools, then calculated $\pi_{diff}$ for each population ($\pi_{diff} = \pi_{naive} - \pi_{recovering}$).
We concatenated the values for *F~ST~* and $\pi_{diff}$ in order of size-sorted chromosome number and adjusted the SNP position based on the relative position in the genome.
We then used the GenWin R package [@beissinger2015] to conduct a splined discrete window analysis for *F~ST~* and $\pi_{diff}$.
This method calculates where non-overlapping window boundaries should occur by identifying inflection points in the fitted spline, therefore balancing false positive and false negative results that occur using other window-based methods [@beissinger2015].
This method also calculates a W-statistic allowing for outlier identification.
We identified outliers as those with a W-statistic greater than 4 standard deviations above the mean for *F~ST~* or above/below the mean for $\pi_{diff}$.
Shared outliers were then identified as those that were outliers in both analyses, meaning that they showed (1) high differentiation between naive and recovering populations, and (2) differential patterns of nucleotide diversity in the same region.
Finally, we extracted gene transcripts mapped within each region and retrieved annotation for that region using BLAST (SI: datasets S3, S4).

Using the liberal set of variant outliers, we performed a statistical overrepresentation test (Fisher's exact test) using Blast2GO [@conesa2005].
First, we retrieved the BLAST hits and mapped GO terms for each gene in our targeted transcriptome.
Next, we compared the 30 <!--# Allie - Check for accuracy --> unique outlier genes from the liberal set of outlier variants to the complete set of genes in our target transcriptome to determine if any GO biological functions, molecular functions, or cellular processes were overrepresented.
We then repeated this process for the set of 33 genes located in the overlapping *F~ST~* and $\pi_{diff}$ splined windows.
Code used for genomic analyses and to create figures is available from the following GitHub repository: <https://github.com/allie128/mylf-selection>.
<!--# Make sure that code have been uploaded to this link before manuscript submission. -->

@@

We conducted `r transno_min`--`r transno_max` translocations per site (@fig-yosemap) and each translocated cohort included `r transfrogs_min` to `r transfrogs_max` frogs (median = `r transfrogs_median`).
In preparation for each translocation, adult frogs were collected from the donor population and measured, weighed, swabbed, and PIT tagged.
Frogs were transported to the recipient site either on foot or via helicopter.
Following release, translocated populations were visited approximately once per month during the summer active season (median number of visits per summer = `r round(surveys_yr_median)`; range = `r surveys_yr_min`--`r surveys_yr_max`) and assessed using diurnal CMR surveys and VES.
CMR surveys allowed esimation of adult survival, adult population size, and recruitment of new adults, and VES provided estimates of tadpole and juvenile abundance.
During 2006-2012, CMR surveys were conducted on a single day (primary period) per site visit, during which all habitats were searched repeatedly for adult frogs.
Frogs were captured using handheld nets, identified via their PIT tag (or tagged if they were untagged), measured, weighed, swabbed, and released at the capture location.
During 2013-2022, we generally used a robust design in which all habitats were searched during several consecutive days (median number of secondary periods per primary period = `r round(robust_year_median)`; range = `r robust_year_min`--`r robust_year_max`), and frogs were processed as described above.
However, when the number of frogs detected on the first survey day was zero or near zero, we typically conducted only a single-day CMR survey.
When using a robust design, within a primary period, frogs that were captured during more than one secondary period were measured, weighed, and swabbed during the first capture, and during subsequent captures were only identified and released.
Frog skin swabs were analyzed using standard Bd DNA extraction and qPCR methods.
For details of survey and swab analysis methods, see [@joseph2018].

@@

We analyzed skin swabs using standard Bd DNA extraction and qPCR methods [@boyle2004, see SI for details].
To assess the magnitude of changes in Bd load on frogs following translocation, we compared Bd loads measured before versus after translocation.
Before-translocation loads were quantified using skin swabs collected from all to-be-translocated frogs at the donor site on the day before or the day of the translocation.
After-translocation Bd loads were based on all swabs collected from translocated frogs at the recipient site in the year of and the year following translocation.
Individual frogs and their associated Bd loads were included in the dataset only if frogs were captured at the recipient site at least once during the year 0-1 period.
We used an analysis workflow similar to that described above.
Briefly, the response variable was before-translocation load and the predictor variable was after-translocation load.
We compared models with and without group-level effects, and group-level effects included pit_tag_id, pit_tag_id nested within site_id, or pit_tag_id nested within translocation_id.
All models used a negative binomial family.
Priors, chains, iterations, and checks were as described above.
Code to create the dataset and fit models is available from the following GitHub repository: <https://github.com/SNARL1/translocation>.
<!--# Does it make more sense to only analyze the model with pit_tag_id as the group-level effect? That is the model that is most analogous to a paired t-test. -->

@@

In this model we ignore density-dependent recruitment because we were interested in the growth of the population from an initial reintroduction and whether this growth was sufficient to prevent extinction over 50 years following the introduction.
We also did not directly consider the dynamics of Bd in this model.
We made this decision because (i) translocated populations are infected with Bd at high prevalence [@joseph2018], and (ii) host density does not seem to play a significant role in multi-year Bd infection dynamics in this system [@wilber2022].
Thus, ignoring Bd infection dynamics and instead assuming all host vital rates are in the presence of high Bd prevalence significantly simplifies the model without much loss of realism.
(Additional details regarding incorporating yearly variability in vital rates and estimating model parameters are provided in the SI.)

>>>>

@@

@@

Roland A. Knapp^a,b,^\*, Mark Q. Wilber^c,1^, Allison Q. Byrne^d,e,1^, Maxwell B. Joseph^f,g^, Thomas C. Smith^a,b^, Andrew P. Rothstein^d,e^, Rob L. Grasso^h^, Erica Bree Rosenblum^d,e^

@@

^a^Sierra Nevada Aquatic Research Laboratory, University of California, Mammoth Lakes, CA, 93546

^b^Earth Research Institute, University of California, Santa Barbara, CA, 93106-3060

^c^School of Natural Resources, University of Tennessee Institute of Agriculture, Knoxville, TN, 37996

^d^Department of Environmental Science, Policy, and Management, University of California - Berkeley, Berkeley, CA, 94720-3114

^e^Museum of Vertebrate Zoology, University of California - Berkeley, Berkeley, CA, 94720-3160

^f^Earth Lab, University of Colorado, Boulder, CO, 80303

^g^Current affiliation: Planet, San Francisco, CA, 94107

^h^Resources Management and Science, Yosemite National Park, El Portal, CA, 95318

@@

^1^These authors contributed equally to this work.

<!--#  Preprint Server: bioRxiv  -->

@@

**Key words:** evolutionary rescue, amphibians, population recovery, reintroduction, disease, *Batrachochytrium dendrobatidis*

@@

Understanding how species persist despite accelerating global change is critical for the conservation of biodiversity.
One possibility is rapid evolution that restores positive growth to declining populations ("evolutionary rescue").
Individuals from these rescued populations could also be used to reestablish extinct populations.
We evaluated both possibilities in a once-common frog species driven to the brink of extinction by a disease that has devastated global amphibian biodiversity and for which few options exist to reverse disease impacts.
We found that frogs have evolved increased resistance to infection, facilitating population recovery.
In addition, introduction of frogs from these populations allowed reestablishment despite the ongoing presence of disease.
Such reintroductions could play an important role in biodiversity conservation in our rapidly changing world.

@@

Human activities are intensifying threats to global biodiversity [@ceballos2015], with important implications for ecosystem resilience and human welfare [@naeem2009].
To persist, affected taxa will often need to adapt to their altered environment or face decline and possible extinction.
The process whereby rapid evolutionary change increases the frequency of adaptive alleles and restores positive population growth is known as "evolutionary rescue" [@carlson2014].
In addition to preventing extinction of declining populations, individuals from rescued populations could also disperse, or be reintroduced, to empty habitats and promote larger-scale recovery.
Although this overall scenario is potentially of critical importance for the conservation of an ever-growing number of threatened taxa, descriptions of its occurrence and effectiveness in the wild are lacking.

One consequence of human alteration of the biosphere is an increase in emerging infectious diseases [@jones2008; @fisher2012].
Such diseases pose a severe threat to wildlife populations [@daszak2000], and have caused dramatic declines and extinctions in a wide range of taxa, including echinoderms, mammals, birds, and amphibians [@hewson2014; @samuel2015; @scheele2019; @cunningham2021].
Amid this global loss of biodiversity, some populations decimated by disease have subsequently recovered to varying extents [@newell2013; @voyles2018; @knapp2016].
Although we know relatively little about the mechanisms underlying recovery [@brannelly2021; @russell2020], host evolution is one possibility [@carlson2014; @searle2020].
This process is known from a diversity of wildlife populations [@savage2016; @epstein2016; @gignoux-wolfsohn2021; @holland2022], and often results in increased host resistance and/or tolerance (see [@medzhitov2012] for definitions).
However, linking these evolutionary responses to population recoveries is difficult and few clear examples exist.
In addition to promoting recovery of disease-impacted populations, the resistant/tolerant individuals in such populations could potentially be used to reestablish populations that were extirpated following disease emergence [@mendelson2019; @brannelly2021].
To date, the use of resistant/tolerant individuals for species recovery is rare [but see @joseph2018], and studies that evaluate its potential utility are critically needed.

@@

In the current study, we combine results from genomic analyses, a 15-year landscape-scale frog reintroduction effort, and a detailed population model to provide an unprecedented example of how evolution in response to disease allows the successful reestablishment of extirpated MYL frog populations.
Specifically, we used exome capture methods to show that MYL frogs have undergone substantial evolutionary change associated with Bd exposure, changes that might provide a genetic basis for the increased resistance/tolerance to Bd infection observed in naturally recovering populations [@knapp2016].
We then use the reintroduction of resistant/tolerant frogs to 24 sites and results from long-term capture-mark-recapture (CMR) surveys to describe the dynamics and drivers of population establishment.
These data are also used to parameterize a population model that extends our inferences of population establishment well beyond the temporal extent of the CMR data.
Collectively, these results provide one of the first examples of how evolutionary rescue and the reintroduction of resistant/tolerant animals can allow the large-scale recovery of disease-impacted species.

@@

To determine whether MYL frog populations show an evolutionary response to Bd, we compared frog genomes between populations with contrasting histories of Bd exposure.
Specifically, we compared frog genomes sampled in four populations that have not yet experienced a Bd-caused epizootic ("naive") [@zhou2015] versus in five populations that experienced a Bd epizootic during the past several decades and have since recovered to varying degrees ("recovering"; @fig-allelefrequencies) [@knapp2016; @vredenburg2010].
These recovering populations exist in an enzootic state [@briggs2010], characterized by high Bd prevalence and, in adults, moderate Bd infection intensity ("load") that is typically well below the level expected to cause mortality [@vredenburg2010].
(See SI Appendix for additional details regarding the study design.)

@@

Results from the individual variant and splined window analyses show that recovering populations have signatures of strong selection on multiple regions of the genome.
The analysis of individual variants identified 11 "outlier" SNPs (i.e., showing significantly different allele frequencies between naive versus recovering populations) from 7 distinct genes across 4 contigs (@fig-selectionresults B, C).
One of the 7 identified genes (LOC108802036) did not have an associated annotation.
For the outlier SNPs, frequency differences between the naive and recovering populations ranged from 41% to 86%.
Most of these SNPs showed frequency differences in only a subset of the sampled populations (@fig-allelefrequencies A, B), but the SNP in the RIN3 gene showed consistent differences in frequencies across all populations (@fig-allelefrequencies C).
This is suggestive of strong - and parallel - selection at this locus associated with exposure of frogs to Bd across multiple populations.
(See SI Appendix for additional description of these results.)

The splined window analysis identified 33 outlier regions for $\pi_{diff}$ and 58 outlier regions for *F~ST~* (@fig-spline-manhattan A, B).
Of these, 9 regions were outliers for both metrics ("shared regions") and 2 of these shared regions also contained one or more outlier SNPs (described above).
A total of 33 annotated genes were found in the 9 shared regions (SI Appendix: Datasets S3, S4).
The largest $\pi_{diff}$, indicative of directional selection, occurred in a 163kb region on Contig19, 12.9Mb upstream of the RIN3 outlier SNP (@fig-spline-manhattan C).
This region contains approximately 500 SNPs and one annotated gene called "interferon-induced very large GTPase 1-like" (GVINP1).
Additionally, a shared outlier region on Contig1 contained two complement factor genes (C6 and C7).
Interestingly, this region had a large negative $\pi_{diff}$ value, consistent with balancing selection.
Finally, one shared outlier region on Contig8 contained one outlier SNP (TCF19) and was within 360kb of another outlier SNP (VARS) (@fig-spline-manhattan D, @fig-synteny-plot).
This region (854kb from the beginning of the outlier window to the VARS SNP) contained a total of 8 annotated genes.
In *Xenopus*, five of these genes occur in the extended major histocompatibility complex (MHC) Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the extended MHC Class III region (HSP70, LSM2, VARS) [@ohta2006].
Therefore, this region under selection is part of the extended MHC Class I and III complex and shows synteny with other amphibian genomes.

@@

Estimates of frog survival during the 1-year period following translocation indicate that survival was highly variable between recipient sites, but relatively constant within recipient sites (for the subset of sites that received multiple translocations; @fig-translocation-survival).
As such, the average survival of frogs in the first translocation to a site was an important predictor of survival of frogs in subsequent translocations to that site (@fig-transsurvival-postdens).
This suggests an important effect of site characteristics and/or (site \* frog) interactions on frog survival.
In addition, 1-year survival was higher for frogs translocated later in the study period than earlier (@tbl-survival-earlylate), which we suggest resulted primarily from our improved ability to choose recipient sites with higher habitat quality for *R. sierrae*.
Finally, Bd loads were fairly consistent before versus after translocation and loads were nearly always well below the level indicative of severe chytridiomycosis and associated frog mortality (@fig-bdload-beforeafter, see SI Appendix text for additional details) [@joseph2018; @vredenburg2010].

@@

Disease-induced population declines are decimating global biodiversity [@daszak2000], but strategies to recover affected species are generally lacking.
Here, we tested the possibility that populations of resistant/tolerant individuals could be used to reestablish extirpated populations.
Although previously proposed [@mendelson2019; @scheele2021], this strategy has rarely been used in the wild (see [@joseph2018] for a small-scale example).
In particular, although numerous studies have documented signatures of selection in hosts following disease-induced declines [e.g., @epstein2016; @gignoux-wolfsohn2021], we are unaware of any that used individuals bearing the resulting adaptive alleles to reestablish extirpated populations.
The results of our study provide evidence that (1) MYL frogs have experienced strong selection following exposure to a recently-emerged lethal fungal disease and that some of the genomic regions under selection are associated with genes related to disease resistance/tolerance, and (2) reintroduction of individuals from these resistant/tolerant populations allows the successful reestablishment of viable frog populations despite ongoing disease.

@@

Results from our genomic analyses strongly suggest that natural selection for adaptive alleles and resulting increased resistance/tolerance is at least partially responsible for the reduced susceptibility of frogs in recovering populations.
We identified specific alleles and genomic regions showing signatures of selection between adjacent Bd-naive and recovering MYL frog populations, consistent with strong selection following Bd exposure.
Importantly, some regions under selection are associated with cellular and immunological mechanisms known to contribute to disease resistance, including in amphibians [@zamudio2020].
For example, the MHC plays an important role in immunity.
In our study, we identified a region that shows evidence of selection in recovering populations and contains eight genes associated with either the MHC Class I or Class III regions.
These results corroborate numerous previous studies linking MHC genes to amphibian resistance/tolerance against Bd [e.g., @savage2011; @bataille2015].
Similarly, the region with the strongest indication of directional selection (as measured by $\pi_{diff}$) contains the interferon-related gene GVINP1.
Several previous studies have found this gene to be differentially expressed during Bd infection [e.g., @grogan2018; @ellison2014] and in populations differing in Bd susceptibility [@grogan2018].
This gene is also strongly linked to disease in salmon, explaining a notable 20% of the resistance phenotype [@robledo2020; @robledo2018].
We also identified a region, characterized by high *F~ST~* and low $\pi_{diff}$, that contained the complement genes C6 and C7.
This could indicate that balancing selection is acting in this region of the genome to favor a diverse set of alleles, as is known for C6 in humans [@soejima2005].
Finally, based on the analysis of individual outlier variants, the RIN3 gene showed a consistent pattern of allele frequency differences across all nine of the frog populations sampled in this study, indicating consistent selection in populations distributed across a wide geographic area.
This gene is associated with inflammation response and in *Xenopus* is highly expressed in wound stumps following amputation of tadpole tails [@fukazawa2009].
Also in this analysis, the outlier variant with the lowest p-value (uncharacterized gene LOC108802036) is located adjacent to a type I interferon gene (Np-IFNi2) in another frog genome [@gan2018], and together with GVINP1 further indicates the importance of interferon-related genes in this system.

@@

To visualize the genomic relationship of our populations we conducted a PCA using the glPCA function in the adegenet R package [@jombart2008].
To detect regions of the genome under selection between naive and recovering populations, we used two approaches: (1) a multivariate linear mixed model to evaluate individual variants (SNPs and INDELs), and (2) a splined window analysis to evaluate larger genomic regions.
For the variant analysis, we first used a stringent data filter to include only variants with \< 5% missing data (missing for no more than 2 individuals), and then calculated the likelihood ratio statistic for the resulting set of 148,307 high quality variants across 127 contigs using GEMMA [@zhou2014].
GEMMA calculates and incorporates a relatedness matrix for input samples, allowing us to account for relatedness and population structure when calculating likelihood ratio statistics.
We identified variants showing different allele frequencies between naive versus recovering populations ("outliers") using a Bonferroni- corrected significance level of 0.01.
We visualized the results using a Manhattan plot and qqplot.
We developed a more liberal set of outlier variants using a Bonferroni- corrected significance level of 0.05 and used this set solely for the gene ontology (GO) analysis (see below and SI Appendix: Dataset S1, S2).

@@

In the splined window analysis, we identified outlier regions using *F~ST~* and differences in nucleotide diversity ($\pi_{diff}$) between naive and recovering populations.
First, we calculated per-site *F~ST~* between the naive and recovering individuals for all bi-allelic SNPs in the 30 largest contigs (98% of all SNPs) using VCFtools [@danecek2011].
Next, we calculated per-site nucleotide diversity $\pi$ separately for individuals from the naive and recovering populations using VCFtools, then calculated $\pi_{diff}$ for each population ($\pi_{diff} = \pi_{naive} - \pi_{recovering}$).
We concatenated the values for *F~ST~* and $\pi_{diff}$ in order of size-sorted chromosome number and adjusted the SNP position based on the relative position in the genome.
We then used the GenWin R package [@beissinger2015] to conduct a splined discrete window analysis for *F~ST~* and $\pi_{diff}$.
This method calculates where non-overlapping window boundaries should occur by identifying inflection points in the fitted spline, therefore balancing false positive and false negative results that occur using other window-based methods [@beissinger2015].
This method also calculates a W-statistic allowing for outlier identification.
We identified outliers as those with a W-statistic greater than 4 standard deviations above the mean for *F~ST~* or above/below the mean for $\pi_{diff}$.
Shared outliers were then identified as those that were outliers in both analyses, meaning that they showed (1) high differentiation between naive and recovering populations, and (2) differential patterns of nucleotide diversity in the same region.
Finally, we extracted gene transcripts mapped within each region and retrieved annotation for that region using BLAST (SI Appendix: Datasets S3, S4).

Using the liberal set of variant outliers, we performed a statistical overrepresentation test (Fisher's exact test) using Blast2GO [@conesa2005].
First, we retrieved the BLAST hits and mapped GO terms for each gene in our targeted transcriptome.
Next, we compared the 30 unique outlier genes from the liberal set of outlier variants to the complete set of genes in our target transcriptome to determine if any GO biological functions, molecular functions, or cellular processes were overrepresented.
We then repeated this process for the set of 33 genes located in the overlapping *F~ST~* and $\pi_{diff}$ splined windows.
Code used for genomic analyses and to create figures is available from the following GitHub repository: <https://github.com/allie128/mylf-selection>.
<!--# Make sure that code has been uploaded to this link before manuscript submission. -->

@@

We conducted `r transno_min`--`r transno_max` translocations per site (@fig-yosemap) and each translocated cohort included `r transfrogs_min` to `r transfrogs_max` frogs (median = `r transfrogs_median`).
In preparation for each translocation, adult frogs were collected from the donor population and measured, weighed, swabbed, and PIT tagged.
Frogs were transported to the recipient site either on foot or via helicopter.
Following release, translocated populations were visited approximately once per month during the summer active season (median number of visits per summer = `r round(surveys_yr_median)`; range = `r surveys_yr_min`--`r surveys_yr_max`) and assessed using diurnal CMR surveys and VES.
CMR surveys allowed estimation of adult survival, adult population size, and recruitment of new adults, and VES provided estimates of tadpole and juvenile abundance.
During 2006-2012, CMR surveys were conducted on a single day (primary period) per site visit, during which all habitats were searched repeatedly for adult frogs.
Frogs were captured using handheld nets, identified via their PIT tag (or tagged if they were untagged), measured, weighed, swabbed, and released at the capture location.
During 2013-2022, we generally used a robust design in which all habitats were searched during several consecutive days (median number of secondary periods per primary period = `r round(robust_year_median)`; range = `r robust_year_min`--`r robust_year_max`), and frogs were processed as described above.
However, when the number of frogs detected on the first survey day was zero or near zero, we typically conducted only a single-day CMR survey.
When using a robust design, within a primary period, frogs that were captured during more than one secondary period were measured, weighed, and swabbed during the first capture, and during subsequent captures were only identified and released.
Frog skin swabs were analyzed using standard Bd DNA extraction and qPCR methods.
For details of survey and swab analysis methods, see [@joseph2018].

@@

We analyzed skin swabs using standard Bd DNA extraction and qPCR methods [@boyle2004, see SI Appendix for details].
To assess the magnitude of changes in Bd load on frogs following translocation, we compared Bd loads measured before versus after translocation.
Before-translocation loads were quantified using skin swabs collected from all to-be-translocated frogs at the donor site on the day before or the day of the translocation.
After-translocation Bd loads were based on all swabs collected from translocated frogs at the recipient site in the year of and the year following translocation.
Individual frogs and their associated Bd loads were included in the dataset only if frogs were captured at the recipient site at least once during the year 0-1 period.
We used an analysis workflow similar to that described above.
Briefly, the response variable was before-translocation load and the predictor variable was after-translocation load.
We compared models with and without group-level effects, and group-level effects included pit_tag_id, pit_tag_id nested within site_id, or pit_tag_id nested within translocation_id.
All models used a negative binomial family.
Priors, chains, iterations, and checks were as described above.
Code to create the dataset and fit models is available from the following GitHub repository: <https://github.com/SNARL1/translocation>.
<!--# Does it make more sense to only analyze the model with pit_tag_id as the group-level effect? That is the model that is most analogous to a paired t-test. -->

@@

In this model we ignore density-dependent recruitment because we were interested in the growth of the population from an initial reintroduction and whether this growth was sufficient to prevent extinction over 50 years following the introduction.
We also did not directly consider the dynamics of Bd in this model.
We made this decision because (i) translocated populations are infected with Bd at high prevalence [@joseph2018], and (ii) host density does not seem to play a significant role in multi-year Bd infection dynamics in this system [@wilber2022].
Thus, ignoring Bd infection dynamics and instead assuming all host vital rates are in the presence of high Bd prevalence significantly simplifies the model without much loss of realism.
(Additional details regarding incorporating yearly variability in vital rates and estimating model parameters are provided in the SI Appendix.)
